---
title: "SOL_ACE"
author: "Yinxian Chen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(nlme)
library(clubSandwich)
#install.packages("devtools")
#devtools::install_github("kupietz/kableExtra")
library(kableExtra)
library(knitr)
library(mice)
library(miceadds)
```

```{r}
sol.pheno<-read.csv("pheno_final_983obs_withcells.csv")
sol.v1.eaa<-read.csv("HCHS_epigen_age_V1_updated_100523.csv")
sol.v2.eaa<-read.csv("HCHS_epigen_age_V2_updated_100523.csv")
sol.v1.dun<-read.csv("HCHS_DunedinPACE_Visit1_updated_100523.csv")
sol.v2.dun<-read.csv("HCHS_DunedinPACE_Visit2_updated_100523.csv")
sol.ac<-read.csv("subject_annotation_2017-05-24.csv")
sol.ace.items<-read.csv("ACE_items.csv")
```

### Data preparation

#### Exposure

  - The primary analysis focused on binary ACEs (<4 ACEs vs. >=4 ACEs), and total ACEs score (continuous)
  
  - In sensitivity analysis, also focused the three-level ACEs (0, 1-3, >=4 ACEs)
  
#### Outcome

  - Only focused on GrimAge and DunedinPACE
  
  - For GrimAge, two indicators were used
  
    - GrimAge acceleration (residuals)
    
    - GrimAge change 
  
  - Only focus on subjects with two Epigenetic age measures

#### Covariates

  - First set (adjusted in Model 1)
  
    - Baseline Age
    
    - Gender
    
    - Study center
    
    - Time variable (AGE_V2-AGE_V1)
    
    - ACEs*Time (To indicate the additional change in rate of EAA by ACEs)
  
  - Second set (further adjusted in Model 2)
  
    - Education attainment (less than high school vs. high school and above)
    
    - Age at arrival in the US (born in the US, arrived before 18y, arrived after 18y)
    
    - Hispanic/Latino background (Central American, Cuba, Dominican, Dominican, Puerto Rican, South American, Mixed/Other)
    
  - For sensitivity analysis:
  
    - Cell types (NK, B, MO, GR, CD4, CD8)
    
    - Ancestry background (five principal components)
    
**Additional information**: One subject's (ID: 1494646) AGE_V1 > AGE_V2. I have flipped the subject's age to the right position.
    
```{r, include=F}
## Data preparation
### Only focus on the GrimAge
sol.v1.grim<-sol.v1.eaa%>%mutate(eaa_grim_v1=AgeAccelGrim,
                                 eage_grim_v1=DNAmGrimAgeBasedOnRealAge,
                                 age_v1=age)%>%
  select(c(ID,eaa_grim_v1,eage_grim_v1,age_v1))


sol.v2.grim<-sol.v2.eaa%>%mutate(eaa_grim_v2=AgeAccelGrim,
                                 eage_grim_v2=DNAmGrimAgeBasedOnRealAge,
                                 age_v2=age)%>%
  select(c(ID,eaa_grim_v2,eage_grim_v2,age_v2))

### Also include the Pace of EAA (DunedinPACE)
sol.v1.dun<-sol.v1.dun%>%mutate(dunedin_v1=DunedinPACE)%>%select(c(ID, dunedin_v1))
sol.v2.dun<-sol.v2.dun%>%mutate(dunedin_v2=DunedinPACE)%>%select(c(ID, dunedin_v2))
sol.ac.pc<-sol.ac%>%select(c(HCHS_ID,EV1,EV2,EV3,EV4,EV5))


### ACEs items

#colnames(sol.ace.items)[1:2]<-c("ID","sub_id")

ace_item<-names(sol.ace.items)[54:63]

##Merge phenotype + epigenetic age variables
sol.list<-list(sol.pheno,sol.ace.items,sol.v1.grim,sol.v2.grim,sol.v1.dun,sol.v2.dun)
sol.phe.eaa<-sol.list%>%reduce(inner_join,by="ID")

# Has corrected
#outlier_ID<-sol.phe.eaa$ID[which((sol.phe.eaa$age_v2-sol.phe.eaa$AGE<3|sol.phe.eaa$age_v2-sol.phe.eaa$AGE>9))]

sol.phe.eaa<-sol.phe.eaa%>%mutate_at(vars(STEA11:STEA20), as.numeric)

#sol.phe.eaa$ACE_TOT_ND<-rowSums(sol.phe.eaa[,c(403:407,409:412)],na.rm = T)

## Create a working dataset
cov_list<-c("ID","AGE_correct","GENDER","PIEA22A","PIEA22B","PIEA23A","PIEA23B","US_BORN","YRSUS","EDUCATION_C2","INCOME_C3","BKGRD1_C7","NK_1","NK_2","B_1","B_2","MO_1","MO_2",
            "GR_1","GR_2","CD4_1","CD4_2","CD8_1","CD8_2","WEIGHT_NORM_OVERALL_EPIGEN")

#cov_list_short<-c("ID","AGE","GENDER","PIEA22A","PIEA22B","PIEA23A","PIEA23B","US_BORN","YRSUS")

#cov_list_tv<-c("ID","NK_1","NK_2","B_1","B_2","MO_1","MO_2",
               #"GR_1","GR_2","CD4_1","CD4_2","CD8_1","CD8_2","WEIGHT_NORM_OVERALL_EPIGEN")

ace.1<-"ACE_TOT"

ace.2<-ace_item

#ace.3<-"ACE_TOT_ND"

eaa<-c("eaa_grim_v1","eaa_grim_v2","eage_grim_v1","eage_grim_v2","age_v1","age_v2","dunedin_v1","dunedin_v2")

sol.ana.1<-sol.phe.eaa[,c(cov_list,ace.1,ace.2,eaa)]

#table(sol.ana.1$PIEA22A)
#table(sol.ana.1$PIEA22B)
# 7TH GRADE 3RD GRADE 4TH GRADE MD MILITARY SCHOOL GED
#table(sol.ana.1$PIEA23A)
#table(sol.ana.1$PIEA23B)
# 4TH GRADE PHD 1ST GRADE 7TH GRADE MASTER DEGREE 3RD GRADE

### Recode the parental education levels

sol.ana.1<-sol.ana.1%>%mutate(fa_edu = ifelse(PIEA22A!=7,PIEA22A,
                                                     ifelse(PIEA22B=="3RD GRADE"|PIEA22B=="4TH GRADE",2,ifelse(PIEA22B=="7TH GRADE",3,
                                                                                                               ifelse(PIEA22B=="MD"|PIEA22B=="MILITARY SCHOOL",6,
                                                                                                                      ifelse(PIEA22B=="GED",4,NA))))))

sol.ana.1<-sol.ana.1%>%mutate(ma_edu = ifelse(PIEA23A!=7,PIEA23A,
                                                     ifelse(PIEA23B=="1ST GRADE"|PIEA23B=="3RD GRADE"|PIEA23B=="4TH GRADE",2,ifelse(PIEA23B=="7TH GRADE",3,
                                                                                                               ifelse(PIEA23B=="PHD"|PIEA23B=="MASTER DEGREE",6,NA)))))


sol.ana.1<-sol.ana.1%>%select(-c(PIEA22B,PIEA23B,PIEA22A,PIEA23A))%>%mutate(ma_edu_3c = factor(ifelse(ma_edu<3,1,
                                                       ifelse(ma_edu==3,2,3))),
                                    fa_edu_3c = factor(ifelse(fa_edu<3,1,
                                                       ifelse(fa_edu==3,2,3))))%>%select(-c(ma_edu,fa_edu))

sol.ana.1$age_v1<-sol.ana.1$AGE_correct
#sol.ana.1<-sol.ana.1%>%select(-c(PIEA22A,PIEA23A))


missing_rate<-do.call("cbind",lapply(sol.ana.1, function(x)
  rbind(sum(is.na(x)),100*sum(is.na(x)/dim(sol.ana.1)[1]))))
colnames(missing_rate)<-names(sol.ana.1)
round(missing_rate, 2)

missing_ace<-sol.ana.1%>%filter(is.na(ACE_TOT))%>%select(c(STEA11:STEA20))

#sol.ana.1<-sol.ana.1%>%filter(!is.na(ACE_TOT))


#summary(sol.ana.1$ACE_TOT_ND)
```

```{r, include=F}
### Convert into the long format for longitudinal analysis
sol.ana.long<-reshape(data = sol.ana.1,
                      varying = list(c("eaa_grim_v1","eaa_grim_v2"),c("eage_grim_v1","eage_grim_v2"),
                                     c("age_v1","age_v2"), c("dunedin_v1","dunedin_v2"),c("NK_1","NK_2"),
                                     c("B_1","B_2"),c("MO_1","MO_2"),c("GR_1","GR_2"),
                                     c("CD4_1","CD4_2"),c("CD8_1","CD8_2")),
                      v.names = c("eaa_grim","eage_grim","age_t","dunedin","NK","B","MO",
                                  "GR","CD4","CD8"),
                      idvar = "ID",
                      timevar = "time",
                      times = c(0,6),
                      direction = "long")
sol.ana.long<-sol.ana.long[order(sol.ana.long$ID),]
#sol.ana.long$t<-rep(c(1:2),nrow(sol.ana.1))
#sol.ana.long$time<-factor(sol.ana.long$time)
sol.ana.long<-sol.ana.long%>%mutate(#age_arr_US = factor(ifelse(US_BORN==0|is.na(US_BORN),ifelse(AGE-YRSUS<18,2,3),1)),
                                    age_change = age_t-AGE_correct,
                                    GENDER = factor(ifelse(GENDER=="M",0,1)))%>%select(-c(age_t,time))

```

```{r, echo=F}
result.summary<-function(data, exposure, covar="normal") {
  
  exposure<-exposure
  covar<-covar
  
  if(exposure =="ace_c4"){
    gee.summary.1<-matrix(NA,nrow = 10, ncol = 4)
    for(i in c(1:10)) {
      indices<-c(2,3,4,10,11,2,3,4,19,20)
      if (i<=5) {
        gee.summary.1[i,1]<-data[[1]]$beta[indices[i]]
        gee.summary.1[i,2]<-data[[1]]$beta[indices[i]]-1.96*data[[1]]$SE[indices[i]]
        gee.summary.1[i,3]<-data[[1]]$beta[indices[i]]+1.96*data[[1]]$SE[indices[i]]
        gee.summary.1[i,4]<-data[[1]]$p_z[indices[i]]
      }
      
      else {
        gee.summary.1[i,1]<-data[[2]]$beta[indices[i]]
        gee.summary.1[i,2]<-data[[2]]$beta[indices[i]]-1.96*data[[2]]$SE[indices[i]]
        gee.summary.1[i,3]<-data[[2]]$beta[indices[i]]+1.96*data[[2]]$SE[indices[i]]
        gee.summary.1[i,4]<-data[[2]]$p_z[indices[i]]
      }
    }
    rownames(gee.summary.1)<-rep(c("1-3 ACEs",">=4 ACEs","Time","1-3 ACEs*Time",">=4 ACEs*Time"),2)
  }
  
  else{
    if(covar=="normal") {
      gee.summary.1<-matrix(NA,nrow = 6, ncol = 4)
      for (i in c(1:6)) {
        indices<-c(2,3,9,2,3,18)
        if (i<=3) {
          
          gee.summary.1[i,1]<-data[[1]]$beta[indices[i]]
          gee.summary.1[i,2]<-data[[1]]$beta[indices[i]]-1.96*data[[1]]$SE[indices[i]]
          gee.summary.1[i,3]<-data[[1]]$beta[indices[i]]+1.96*data[[1]]$SE[indices[i]]
          gee.summary.1[i,4]<-data[[1]]$p_z[indices[i]]
        }
        
        else {
          gee.summary.1[i,1]<-data[[2]]$beta[indices[i]]
          gee.summary.1[i,2]<-data[[2]]$beta[indices[i]]-1.96*data[[2]]$SE[indices[i]]
          gee.summary.1[i,3]<-data[[2]]$beta[indices[i]]+1.96*data[[2]]$SE[indices[i]]
          gee.summary.1[i,4]<-data[[2]]$p_z[indices[i]]
        }
      }
      
      if(exposure=="ace_c2"){
        rownames(gee.summary.1)<-rep(c(">=4 ACEs", "Time", ">=4 ACEs*Time"),2)
      }
      else {
        rownames(gee.summary.1)<-rep(c("ACEs total score", "Time", "ACEs total score*Time"),2)
      }
    }
    
    if(covar=="ancenstry") {
      gee.summary.1<-matrix(NA,nrow = 3, ncol = 4)
      for(i in c(1:3)) {
        indices<-c(2,3,23)
        gee.summary.1[i,1]<-data$beta[indices[i]]
        gee.summary.1[i,2]<-data$beta[indices[i]]-1.96*data$SE[indices[i]]
        gee.summary.1[i,3]<-data$beta[indices[i]]+1.96*data$SE[indices[i]]
        gee.summary.1[i,4]<-data$p_z[indices[i]]
      }
      if(exposure=="ace_c2"){
        rownames(gee.summary.1)<-c(">=4 ACEs", "Time", ">=4 ACEs*Time")
      }
      else {
        rownames(gee.summary.1)<-c("ACEs total score", "Time", "ACEs total score*Time")
      }
    }
    
    if(covar=="cell_type") {
      gee.summary.1<-matrix(NA,nrow = 3, ncol = 4)
      for(i in c(1:3)) {
        indices<-c(2,3,24)
        gee.summary.1[i,1]<-data$beta[indices[i]]
        gee.summary.1[i,2]<-data$beta[indices[i]]-1.96*data$SE[indices[i]]
        gee.summary.1[i,3]<-data$beta[indices[i]]+1.96*data$SE[indices[i]]
        gee.summary.1[i,4]<-data$p_z[indices[i]]
      }
      if(exposure=="ace_c2"){
        rownames(gee.summary.1)<-c(">=4 ACEs", "Time", ">=4 ACEs*Time")
      }
      else {
        rownames(gee.summary.1)<-c("ACEs total score", "Time", "ACEs total score*Time")
      }
    }
  }
  
  colnames(gee.summary.1)<-c("beta","ll","ul","p-value")
  return(gee.summary.1)
}
```

### Longitudinal analysis setup

  - Generalized estimating equations (GEE) and Generalized linear mixed-effect models (GLMM)
  
  - Each model can be weighted by the the sampling weight for each subject
  
  - Advantages to use the longitudinal analysis approach:
  
    - Can gain the estimates of associations in interest simultaneously:
    
      - Association between ACEs and the EAA at baseline
      
      - Association between ACEs and the rate of change in EAA from time 1 to time 2
      
    - More powerful than modeling two outcomes separately and can also have a consistent interpretation
    
  - Robust variance estimator (sandwish estimator) was used to account for the sampling weight

#### Imbalanced desgin

  - We can see that not all participants were measured the second epigenetic age after 6 years from the first visit
  
  - Also, there were some outliers with unreasonable interval between two visits
  
    - First visit (2008-2011) and Second visit (2014-2017)
    
      - So the range of the interval should be [2014-2011, 2017-2008] = [3, 9] years
  
    - I removed those out of this range in the analysis but will definitely add them back if necessary
  
  - The GEE cannot handle the imbalanced design in the longitudinal study
  
    - So only GLMM could be used here

```{r}
# Consistency between baseline age and age at first epigenetic age measurement
summary(sol.ana.1$age_v1-sol.ana.1$AGE)
# Interval between first and second measurement
table(sol.ana.1$age_v2-sol.ana.1$AGE)
# Subjects that had the interval <3 or >9
sol.ana.1$ID[which((sol.ana.1$age_v2-sol.ana.1$AGE<3|sol.ana.1$age_v2-sol.ana.1$AGE>9))]
outlier_ID<-sol.ana.1$ID[which((sol.ana.1$age_v2-sol.ana.1$AGE<3|sol.ana.1$age_v2-sol.ana.1$AGE>9))]

```

#### Create separate dataset for different ACEs indicators
```{r}
### Primary analysis (total score)
sol.ana.long.1<-sol.ana.long[,!names(sol.ana.long)%in%c(ace.1)]

### Sensitivity 1 (each type of ACE)

### Sensitivity 2 (parental separate as covariate)
```


### Test for multiple imputation analysis

```{r}
predma<-quickpred(sol.ana.long.1, exclude = c("ID","WEIGHT_NORM_OVERALL_EPIGEN"))
imp_1<-mice(sol.ana.long.1, m=1, predictorMatrix = predma, print = F)
#summary(sol.ana.long.1)

#imp_1$predictorMatrix
#predma
predmethod<-imp_1$method
#predmethod
#summary(complete(imp_1))

lambda <- 1- sum(complete.cases(sol.ana.long.1))/dim(sol.ana.long.1)[1]
RE <- 0.95
lambda/(RE^(-1)-1) #5

imp_5<-mice(sol.ana.long.1, m=5, maxit = 10, predictorMatrix = predma, method = predmethod, print = F)

plot(imp_5)

imp_list<-datlist_create(imp_5)

```


### Primary analysis

  - Although statistically insignificant in all Model 2 (adjusted for the proximal childhood SES)

    - Stronger evidence that two-level ACEs was associated with increased **DunedinPACE** at the first visit in in two models
  
    - Stronger evidence that total ACEs score was associated with increased **GrimAA** and **GrimAage** at the first visit in two models
  
  - Different reference group led to different results
  
    - We can see consistent results in the models using three-level ACEs with models using ACEs score
    
  - What consistent is, **no association** was seen in the rate of three EAA indicators (results shown in the ACE*time)

```{r}
### Unbalance design
glm.ace.ub<-function(data, exposure, outcome) {
  
  formula_1<-as.formula(paste(outcome, "~", exposure,"+ age_change + AGE + GENDER + CENTER + age_change:", exposure))
  model.1<-lme(formula_1, 
               data = get(data),
               random = ~1+age_change|ID,
               weights = varFixed(~WEIGHT_NORM_OVERALL_EPIGEN),
               method = "REML",
               na.action = na.omit)
  #print(summary(model.1))
  robust.1<-coef_test(model.1, vcov = "CR0", test = "z")
  
  
  formula_2<-as.formula(paste(outcome, "~", exposure,"+ age_change + AGE + GENDER + CENTER + EDUCATION_C2 + age_arr_US + BKGRD1_C7 + age_change:", exposure))
  model.2<-lme(formula_2, 
               data = get(data),
               random = ~1+age_change|ID,
               weights = varFixed(~WEIGHT_NORM_OVERALL_EPIGEN),
               method = "REML",
               na.action = na.omit)
 # print(summary(model.2))
  robust.2<-coef_test(model.2, vcov = "CR0", test = "z")
  return(list(robust.1,robust.2))
}
```

```{r, echo=F}
ana.1.ub<-glm.ace.ub("sol.ana.long.ub", "ace_c2", "eaa_grim") #Binary ACE (<4 vs. >=4) and EAA GrimAge
#ana.1.ub
ana.2.ub<-glm.ace.ub("sol.ana.long.ub", "ace_c2", "eage_grim") #Binary ACE (<4 vs. >=4) and Epigenetic age change using GrimAge
#ana.2.ub
ana.3.ub<-glm.ace.ub("sol.ana.long.ub", "ace_c2", "dunedin") #Binary ACE (<4 vs. >=4) and Epigenetic age Pace
#ana.3.ub
ana.4.ub<-glm.ace.ub("sol.ana.long.ub", "ACE_TOT", "eaa_grim") #Continuous ACE and EAA GrimAge
#ana.4.ub
ana.5.ub<-glm.ace.ub("sol.ana.long.ub", "ACE_TOT", "eage_grim") #Continuous ACE and Epigenetic age change using GrimAge
#ana.5.ub
ana.6.ub<-glm.ace.ub("sol.ana.long.ub", "ACE_TOT", "dunedin") #Continuous ACE and Epigenetic age Pace
#ana.6.ub

summary.1.ub<-result.summary(ana.1.ub, exposure = "ace_c2")
summary.2.ub<-result.summary(ana.2.ub, exposure = "ace_c2")
summary.3.ub<-result.summary(ana.3.ub, exposure = "ace_c2")
summary.4.ub<-result.summary(ana.4.ub, exposure = "ACE_TOT")
summary.5.ub<-result.summary(ana.5.ub, exposure = "ACE_TOT")
summary.6.ub<-result.summary(ana.6.ub, exposure = "ACE_TOT")
```

```{r, echo=F}
test.3<-rbind(summary.1.ub,summary.4.ub,summary.2.ub,
              summary.5.ub,summary.3.ub,summary.6.ub)


kable(test.3, format = "latex", booktabs = T, digits = 3, 
      caption = "Association of ACEs with EAA using different ACEs variables and EAA indicators")%>% 
  add_header_above(c(" " = 1, "GLM" = 4))%>% pack_rows(index = c("GrimAA"=12, "GrimAge"=12,"DunedinPACE"=12))%>%
  pack_rows(index = rep(c("Model 1" = 3, "Model 2" = 3),6))
``` 

### Sensitivity analysis

#### Adjusted for cell types

  - Adjusted for cell types seemed increased the power (many associations turned into statistically significant!!)
  
  - Probably because cell types were associated with the outcome but not the exposure (aka precision variables)

```{r}
glm.ace.cell<-function(data, exposure, outcome) {
  formula.cell<-as.formula(paste(outcome, "~", exposure,"+ age_change + AGE + GENDER + CENTER + EDUCATION_C2 + age_arr_US + BKGRD1_C7 + NK + B + MO + GR + CD4 + CD8 + age_change:", exposure))
  model.cell<-lme(formula.cell, 
               data = get(data),
               random = ~1+age_change|ID,
               weights = varFixed(~WEIGHT_NORM_OVERALL_EPIGEN),
               method = "REML",
               na.action = na.omit)
  #print(summary(model.cell))
  robust.cell<-coef_test(model.cell, vcov = "CR0", test = "z")
  return(robust.cell)
}
```

```{r,echo=F}
ana.1.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ace_c2", "eaa_grim")
#ana.1.ml.cell
ana.2.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ace_c2", "eage_grim")
#ana.2.ml.cell
ana.3.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ace_c2", "dunedin")
#ana.3.ml.cell
ana.4.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ACE_TOT", "eaa_grim")
#ana.4.ml.cell
ana.5.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ACE_TOT", "eage_grim")
#ana.5.ml.cell
ana.6.ml.cell<-glm.ace.cell("sol.ana.long.ub", "ACE_TOT", "dunedin")
#ana.6.ml.cell

summary.1.ml.cell<-result.summary(ana.1.ml.cell, "ace_c2", "cell_type")
summary.2.ml.cell<-result.summary(ana.2.ml.cell, "ace_c2", "cell_type")
summary.3.ml.cell<-result.summary(ana.3.ml.cell, "ace_c2", "cell_type")
summary.4.ml.cell<-result.summary(ana.4.ml.cell, "ACE_TOT","cell_type")
summary.5.ml.cell<-result.summary(ana.5.ml.cell, "ACE_TOT","cell_type")
summary.6.ml.cell<-result.summary(ana.6.ml.cell, "ACE_TOT","cell_type")
```


```{r, echo=F}
test.1<-rbind(summary.1.ml.cell,
            summary.4.ml.cell,
            summary.2.ml.cell,
            summary.5.ml.cell,
            summary.3.ml.cell,
            summary.6.ml.cell)

kable(test.1, format = "latex", booktabs = T, digits = 3, 
      caption = "Sensitivity Analysis by Adjusting for Cell Types")%>% 
  add_header_above(c(" " = 1, "GLM" = 4))%>% 
  pack_rows(index = c("GrimAA"=6, "GrimAge"=6,"DunedinPACE"=6))%>%
  pack_rows(index = rep(c("Model 2 + cell types" = 3),6))
```

#### Using three-level ACEs

  - When changing the reference group from "<4 ACEs" to "No ACE", the association was better revealed in **GrimAA** and **GrimAge**, but not **DunedinPACE**
  
    - Results were consistent with models using ACEs scores
  
  - There is some evidence that even small number of ACEs can associated with EAA
  
  - Maybe consider it the primary exposure variable?


```{r, echo=F}
#### GLM
ana.1.ml.4c<-glm.ace.ub("sol.ana.long.ub", "ace_c4", "eaa_grim") # EAA GrimAge
#ana.1.ml.4c

ana.2.ml.4c<-glm.ace.ub("sol.ana.long.ub", "ace_c4", "eage_grim") #Epigenetic age change using GrimAge
#ana.2.ml.4c

ana.3.ml.4c<-glm.ace.ub("sol.ana.long.ub", "ace_c4", "dunedin") # Epigenetic age Pace
#ana.3.ml.4c

summary.1.4c.ml<-result.summary(ana.1.ml.4c, "ace_c4")
summary.2.4c.ml<-result.summary(ana.2.ml.4c, "ace_c4")
summary.3.4c.ml<-result.summary(ana.3.ml.4c, "ace_c4")
```



```{r, echo=F}
test.2<-rbind(summary.1.4c.ml,
             summary.2.4c.ml,
             summary.3.4c.ml)

kable(test.2, format = "latex", booktabs = T, digits = 3, caption = "Sensitivity Analaysis by Using Three-levels ACEs")%>% add_header_above(c(" " = 1, "GLM" = 4))%>% pack_rows(index = c("GrimAA"=10, "GrimAge"=10,"DunedinPACE"=10))%>%
  pack_rows(index = rep(c("Model 1" = 5, "Model 2" = 5),3))
```

#### Adjust for ancestry

  - Only 774 subjects had the ancestry information
  
  - Also seemed to improve the power (p-value decreased for ACE-EAA associations)

```{r, echo=F}
colnames(sol.ac.pc)[1]<-"ID"
sol.ana.ac<-merge(sol.ana.1,sol.ac.pc, by = "ID")

## Long format dataset including ancestry
sol.ana.long.ac<-reshape(data = sol.ana.ac,
                      varying = list(c("eaa_grim_v1","eaa_grim_v2"),c("eage_grim_v1","eage_grim_v2"),
                                     c("age_v1","age_v2"), c("dunedin_v1","dunedin_v2"),c("NK_1","NK_2"),
                                     c("B_1","B_2"),c("MO_1","MO_2"),c("GR_1","GR_2"),
                                     c("CD4_1","CD4_2"),c("CD8_1","CD8_2")),
                      v.names = c("eaa_grim","eage_grim","age_t","dunedin","NK","B","MO",
                                  "GR","CD4","CD8"),
                      idvar = "ID",
                      timevar = "time",
                      times = c(0,6),
                      direction = "long")
sol.ana.long.ac<-sol.ana.long.ac[order(sol.ana.long.ac$ID),]
sol.ana.long.ac$t<-rep(c(1:2),nrow(sol.ana.ac))
sol.ana.long.ac$time<-factor(sol.ana.long.ac$time)
sol.ana.long.ac<-sol.ana.long.ac%>%mutate(ace_c2 = ifelse(ACE_TOT<4,0,1),
                                    age_arr_US = factor(ifelse(US_BORN==0|is.na(US_BORN),ifelse(AGE-YRSUS<18,2,3),1)),
                                    BKGRD1_C7 = factor(BKGRD1_C7),
                                    ace_c4 = factor(ifelse(ACE_TOT==0,1,
                                                           ifelse(ACE_TOT<4,2,3))),
                                    age_change = age_t-AGE)
```


```{r}
# Consistency between baseline age and age at first epigenetic age measurement
summary(sol.ana.ac$age_v1-sol.ana.ac$AGE)
# Interval between first and second measurement
table(sol.ana.ac$age_v2-sol.ana.ac$AGE)
# Subjects that had the interval <3 or >9
sol.ana.ac$ID[which((sol.ana.ac$age_v2-sol.ana.ac$AGE<3|sol.ana.ac$age_v2-sol.ana.ac$AGE>9))]
outlier_ID_ac<-sol.ana.ac$ID[which((sol.ana.ac$age_v2-sol.ana.ac$AGE<3|sol.ana.ac$age_v2-sol.ana.ac$AGE>9))]

# Exclude the outlier
sol.ana.long.ac.ub<-sol.ana.long.ac%>%filter(!ID%in%outlier_ID_ac)
```


```{r}
#### GLM
glm.ace.ac<-function(data, exposure, outcome) {
  formula.ac<-as.formula(paste(outcome, "~", exposure,"+ age_change + AGE + GENDER + CENTER + EDUCATION_C2 + age_arr_US + BKGRD1_C7 + EV1 + EV2 + EV3 + EV4 + EV5 + age_change:", exposure))
  model.ac<-lme(formula.ac, 
                  data = get(data),
                  random = ~1+age_change|ID,
                  weights = varFixed(~WEIGHT_NORM_OVERALL_EPIGEN),
                  method = "REML",
                  na.action = na.omit)
  #print(summary(model.ac))
  robust.ac<-coef_test(model.ac, vcov = "CR0", test = "z")
  return(robust.ac)
}
```


```{r, echo=F}
ana.1.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ace_c2", "eaa_grim") #Binary ACE (<4 vs. >=4) and EAA GrimAge
#ana.1.ac
ana.2.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ace_c2", "eage_grim") #Binary ACE (<4 vs. >=4) and Epigenetic age change using GrimAge
#ana.2.ac

ana.3.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ace_c2", "dunedin") #Binary ACE (<4 vs. >=4) and Epigenetic age Pace
#ana.3.ac
ana.4.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ACE_TOT", "eaa_grim") #Continuous ACE and EAA GrimAge
#ana.4.ac
ana.5.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ACE_TOT", "eage_grim") #Continuous ACE and Epigenetic age change using GrimAge
#ana.5.ac
ana.6.ac<-glm.ace.ac("sol.ana.long.ac.ub", "ACE_TOT", "dunedin") #Continuous ACE and Epigenetic age Pace
#ana.6.ac

summary.1.ml.ac<-result.summary(ana.1.ac, "ace_c2", "ancenstry")
summary.2.ml.ac<-result.summary(ana.2.ac, "ace_c2", "ancenstry")
summary.3.ml.ac<-result.summary(ana.3.ac, "ace_c2", "ancenstry")
summary.4.ml.ac<-result.summary(ana.4.ac, "ACE_TOT", "ancenstry")
summary.5.ml.ac<-result.summary(ana.5.ac, "ACE_TOT", "ancenstry")
summary.6.ml.ac<-result.summary(ana.6.ac, "ACE_TOT", "ancenstry")
```

```{r, echo=F}
test.4<-rbind(summary.1.ml.ac,
              summary.4.ml.ac,
              summary.2.ml.ac,
              summary.5.ml.ac,
              summary.3.ml.ac,
              summary.6.ml.ac)

kable(test.4, format = "latex", booktabs = T, digits = 3, caption = "Sensitivity Analysis by Adjusting for Ancestry")%>% add_header_above(c(" " = 1, "GLM" = 4))%>% pack_rows(index = c("GrimAA"=6, "GrimAge"=6,"DunedinPACE"=6))%>%
  pack_rows(index = rep(c( "Model 2 + ancenstry" = 3),6))
```
